---
title: "Investigating Covid-19-Virus-Trend"
output: html_document
date: "2024-11-11"
---

# **This tutorial intends to investigate the Covid-19 Virus Trend in different countries**

#### **Importing necessary libraries**


```{r echo=FALSE}
library(dplyr)
library(ggplot2)
library(data.table)
```



```{r}
#Getting the working directory
getwd()
```


#### **Importing data**

```{r}
df_covid19 = fread("tested_worldwide.csv")
```

#### **Showing a subset  of the data**


```{r}
#6 First rows
head(df_covid19)

```
```{r}
#6 Last rows
tail(df_covid19)
```


#### **Displaying the dimension of the dataset**

```{r}
dim(df_covid19)
#The dataset has 27641 observations and 12 columns
```

#### **Puting the columns names in a variable**

```{r}
df_cols <- colnames(df_covid19)
df_cols
```
#### **Some summary statistics**

```{r}
summary(df_covid19)
```
##### **Regarding the summary statistics, we can say that:**

*Data was collected from the 16th January 2020 to the 09th November 2020.
*Some columns have negative values (active, daily_tested, daily_positive) which should 
require our attention.

#### **Let's see what these records look like**


```{r}
df_negative_values <-df_covid19%>% 
  filter(active <0 | daily_tested <0 | daily_positive <0) %>% 
  arrange(Date,Country_Region)
#There are 340 values in that case, 1.2%
```

#### **For simplicity, let's remove these records to prevent baises**

```{r}
df_test <-df_covid19%>% 
  filter(!(active<0 | daily_tested <0 | daily_positive <0)) %>% 
  #filter(active >=0 & daily_tested >=0 & daily_positive >=0) %>%
  arrange(Date,Country_Region)
```



#### **Exploring null values by columns**



```{r}
df_covid19 %>% is.na() %>% colSums()
```


#### **Giving the purcentages of null values by columns**


```{r}
n = dim(df_covid19)[1]
vec <- df_covid19 %>% is.na() %>% colSums()
round((vec/n)*100,2)
#Columns with no null values: Date, Country_Region,  Province_State
#Columns with more than 30% of null values: active, hospitalized, hospitalizedCur and recovered 
```

#

#
```{r}
library(tibble)
```

```{r}
glimpse(df_covid19)
```
#### **Let's look at the countries in the dataset**

```{r}
df_covid19$Country_Region %>% unique()
# There are 147 countries in total
```

#### ** The regions in the dataset**

```{r}
df_covid19$Province_State %>% unique()
# There are 81 provinces
```
```{r}
#test<-df_covid19 %>% select(Country_Region,Province_State) %>% 
  #arrange(Country_Region) %>% distinct()
```




#### **Let's subset the dataset into  the "All States" provinces label in order to make analyses in country level**

```{r}
df_covid19_cl<-df_covid19 %>% 
  filter(Province_State=="All States") %>% 
  select(-c(Province_State))
```


#### **Let's make analyses only on daily variables**

#### **A description of all variables**


```{r}
colnames(df_covid19_cl)
```

* Date: The date 
* Country_Region: The country names
* positive: The cumulative number of positive cases
* active: The number of actively cases on the day
* hospitalized: The cumulative number of hospitalized cases
* hospitalizedCurr: the daily number of hospitalized cases
* recovered: The cumulative number of recovered cases
* death: The cumulative number of dead cases
* total_tested: The cumulative number of tested cases
* daily_tested: The daily number of tests
* daily_positive: The daily number of positive cases



```{r}
df_covid19_cl_d<-df_covid19_cl %>% 
  select(c(Date,Country_Region,active,hospitalizedCurr,daily_tested,daily_positive))
```






#### **Let's aggregate the columns by countries to have the total number of cases by countries**


```{r}
df_covid19_cl1 <- df_covid19_cl %>% 
  group_by(Country_Region) %>% 
  summarise(active=sum(active,na.rm = T),
            hospitalized=sum(hospitalizedCurr,na.rm = T),
           tested = sum(daily_tested,na.rm = T),
           positive = sum(daily_positive,na.rm = T))
```

#### **The top 15 countries with high number of tests**


```{r}
df_covid19_cl1 %>% 
  arrange(desc(tested)) %>% 
  head(15)
```

